Supplementary information for The transboundary nature of the world’s exploited marine species

### 2020-10-05 Scripts/Manuscript_Draft.Rmd
Set manuscript for thesis

The *Sea Around Us* has reconstructed global fisheries catches from 1950 to 2014, identifying commercial marine fish and invertebrates as well as fishing regions [@Zeller2016]. We used the *Sea Around Us* reconstructed species list to determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs. The total number of species analyzed was 938, representing 67% of identified species by the *Sea Around Us* and accounting for 96.52% of the catch identified at the species’ level. To determine the current distribution of exploitable marine species, we used four data sources of species-distributions: (*i*) occurrence data, (*ii*) an Ensemble Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (Table @\ref(tab:Data_Table)). Each source represents a different method of estimating the distribution of a given species, thus providing a more robust result than an analysis focused on a single data source. Only commercial fished species with data from all four sources were included in our analysis.

```{r Data_Table, eval = F, echo = F}

@@

data.frame(Source,
           Abreviation,
           Main_Method,
           Sources) %>% 
  rename("Main Method"=Main_Method) %>% 
  kable("latex", booktabs = T,
        caption = "Summary of data sources to estimate species' distributions. All data has a ressolution of 0.5 degrees latitude x 0.5 degrees longitude. All data is publically available, see references for details.") %>%
  kable_styling(full_width = T) %>%
  column_spec(2,width = "2cm") %>% #wraps text in column2
  column_spec(3,width = "5cm") #%>% #wraps text in column2
  # write_csv("Table_1.csv")
```

@@

We used catch data as described above to determine the catch trend of each species within an EEZ. Although this method has previously been used to estimate stock status [@Grainger:1996tk], the categories presented here are intended to represent catch trends, and not fishing status for each species as many environmental and social-economic factors (e.g., temperature, markets, fishing policies, and fishing effort) affect catches [@Pauly:2013gs; @Branch:2008kk]. We only assessed species within each EEZ for which at least ten years of data were available between 1951 and 2014 and with at least five consecutive years of data. Three final categories were drawn for each species depending on catch volume within each EEZ (e.g., present, maximum, and minimum EEZ's historical catch) and the year (e.g., year of maximum historical catch of the species within that EEZ) (Table Table \@ref(tab:CallTab2)) [@Kleisner:2011wn]. Accordingly, Category A represents fisheries that are registering increases in catch (“increasing”); Category B, species that have a constant catch rate (“constant”); and Category C, species that have registered declines in catch over the last 10 years (“decreasing”). Finally, we reported the predominant category over the time period considered.

@@

data.frame(Categories,
           Rules) %>% 
  kable("latex", 
        booktabs = T,
        caption = "Rules to determine the category of each transboundary species") %>%
  kable_styling(full_width = T) %>%
  column_spec(1,width = "3cm") %>% #wraps text in column2
  column_spec(2,width = "13cm") #%>% #wraps text in column2
  # write_csv("Table_2.csv")

@@

```{r Figure1, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species and their contribution to global fisheries catch and revenue. a) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. b) Contribution of transboundary species to regional revenue (left) and catch (right). Regions classified according to the United Nations sub-regions. Points = mean +- sd. Revenue in 2010 real USD"}

@@

<!-- ![Figure 1. Number of transboundary species and their contribution to global fisheries catch and revenue. a) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. b) Contribution of transboundary species to regional revenue (left) and catch (right). Regions classified according to the United Nations sub-regions: E. = Eastern, N. = Northern, S. = Southern, W. = Western, Ltn. Ame. & the Car. = Latin America and the Caribbean. Points = regional  mean ± sd. Revenue in 2019 real USD. Figure created using R version 3.5.2. Land and EEZ shapefiles from Natural Earth (www.naturalearthdata.com) and Sea Around Us (www.seaaroundus.org), respectively.](../Figures/Fig1.png) -->

@@

Peru and Russia, having the two largest fisheries by total catch in the world [@FAO:2018KK], each responsible for over 5.8 million tonnes annually, on average, between 2005 and 2014  (Supplementary Fig. S1). Peru's catches consist mainly of Peruvian anchoveta (*Engraulis ringens*) and account for 79% of the national transboundary species production. Peru and Chile recently signed an agreement to work towards standardized stock assessments through coordinated management of the southern anchoveta stock [@UNDP:2016wp]. A similar management agreement was signed by Russia, Japan and the USA over shared Alaskan pollock (*Theragra chalcogramma*) in the Bering Sea in 1988 [@NOAA:gX6CtUTk]. Transboundary species also make large contributions to fisheries in Eastern Asia (one-way ANOVA; *Df*(16,158) = 2.265, *p* = 0.005, $\alpha$ = 0.05). China, the world's top fish producer [@FAO:2018KK], obtains one third (5.1 million tonnes) of its total catches from transboundary species, followed by Japan (1.8 million tonnes) and South Korea (1.06 million tonnes). Differences in the regional importance of transboundary fisheries can also be illustrated in terms of catch-revenue over the area ($km^2$) of the EEZ (Figure \@ref(fig:Figure2)). As an example, the aggregated EEZ area for all Northern European countries where transboundary species are present was 3.3 million $km^2$, the $6^{th}$ smallest of the 17 groups analyzed (Supplementary Table S1). However, it had the second highest average revenue (USD 26.1 thousand per $km^2$) and the highest average catch (19.9 tonnes per $km^2$) of transboundary species per EEZ area between 2005 and 2014. At the country level, countries from Western Europe accrued significantly more revenue from transboundary fisheries per $km^2$ than any other country (one-way ANOVA, *Df*(16,165) = 3.267, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.05; Figure \@ref(fig:Figure2)).

```{r Figure_2, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap=""}

@@

<!-- ![Figure 2. Weighted benefits of transboundary species by $km^{2}$ and UN sub-region. a) Revenue in thousand 2019 USD. b) Catch in tonnes. Points are sub-region averages $\pm$ s.d. by country. Figure created using R version 3.5.2. Abbreviations as in Figure 1.](../Figures/Fig2.png) -->

We determined the catch trend of each species within each EEZ, classifying them as increasing (Category A), constant (Category B) or decreasing (Category C) (Figure \@ref(fig:Figure3)). While previous studies have demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Our intention here is to examine where the catch trends of transboundary species differ from non-transboundary species between 2005 and 2010 relative to historic catch since 1951 (See methods). We found significant differences for all catch trend categories for transboundary species (one-way ANOVA, *Df*(2,459) = 47.94, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.001), and no significant differences in catch trend categories for discrete species (one-way ANOVA, *Df*(2,106) = 1.885, *p* = 0.157; $\alpha$ = 0.05). We also found significant differences in catch trends when directly comparing transboundary to discrete species categories (MANOVA, *Df*(2,459) = 19.001, *p* < 0.001). Overall, transboundary species only targeted by one country are generally less likely to have a decreasing catch trend compared to instances where the shared species is fished by neighboring countries (Supplementary Table S2).

```{r FIgure_3, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species by catch category. Category A, increasing catch; Category B, constant catch; Category C, decreasing catch"}

@@

```{r Figure_4, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of countries shared by transboundary species. Showing only species that share > 20 countries"}

@@

```{r CallSup1, eval = T, fig.width=6, fig.height=4, echo=FALSE, fig.cap="Number of transboundary species per EEZ and they contribution to countries' EEZ catch. The number of species is represented in the EEZ while the catch (thousand tons) is represented in the land polygon."}

img <- readPNG("../Figures/FigS1.png")
grid.raster(img)
```

<!-- ![Figure S1. Number of transboundary species per EEZ and their contributions to countries’ EEZ catch. The number of species is displayed in the EEZ while catch (in thousand tons) is displayed in the land polygon.](~/GitHub/FishForVisa/Figures/FigS1.png) -->

The number of gridcells in which a species is present (*Area Index*) is expected to directly influence whether or not a species is considered transboundary. Therefore, we tested the sensitivity of our results to the *Area Index* by estimating the number of transboundary species along a gradient of values (Figure \@(fig:CallSup2)). As expected, using a more relaxed value (e.g. the species is present within 10% or more of the EEZ) will result in a greater number of transboundary species (median of 40 EEZs shared per species), while a more restrictive value (e.g., 50%) results in a median estimate of 4. The current analysis utilized a 25% *Area Index* threshold (Figure \@ref(fig:CallSup2) darker histogram).

```{r CallSup2, eval = T, fig.width=6, fig.height=4, echo=FALSE, fig.cap="Histogram of transboundary species using different values of Area Index. Darker histogram represents used value (25%) in the analysis"}

img <- readPNG("../Figures/FigS2.png")
grid.raster(img)
```

<!-- ![Figure S2. Histogram of number of transboundary species using different Area Index threshold values. The darker histogram represents the threshold (25%) used in the present analysis](~/GitHub/FishForVisa/Figures/FigS2.png) -->

>>>>

The *Sea Around Us* has reconstructed global fisheries catches from 1950 to 2014, identifying commercial marine fish and invertebrates as well as fishing regions [@Zeller2016]. We used the *Sea Around Us* reconstructed species list to determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs. The total number of species analyzed was 938, representing 67% of identified species by the *Sea Around Us* and accounting for 96.52% of the catch identified at the species’ level. To determine the current distribution of exploitable marine species, we used four data sources of species-distributions: (*i*) occurrence data, (*ii*) an Ensemble Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (Table @\ref(tab:CallTab1)). Each source represents a different method of estimating the distribution of a given species, thus providing a more robust result than an analysis focused on a single data source. Only commercial fished species with data from all four sources were included in our analysis.

```{r CallTab1, eval = T, echo = F}

@@

data.frame(Source,
           Abreviation,
           Main_Method,
           Sources) %>% 
  rename("Main Method"=Main_Method) %>% 
  kable("latex", 
        booktabs = T,
        caption = "Summary of data sources to estimate species' distributions. All data has a ressolution of 0.5 degrees latitude x 0.5 degrees longitude. All data is publically available, see references for details") %>%
  kable_styling(full_width = T) %>%
  column_spec(2,width = "2cm") %>% #wraps text in column2
  column_spec(3,width = "5cm") # #wraps text in column2
```

@@

We used catch data as described above to determine the catch trend of each species within an EEZ. Although this method has previously been used to estimate stock status [@Grainger:1996tk], the categories presented here are intended to represent catch trends, and not fishing status for each species as many environmental and social-economic factors (e.g., temperature, markets, fishing policies, and fishing effort) affect catches [@Pauly:2013gs; @Branch:2008kk]. We only assessed species within each EEZ for which at least ten years of data were available between 1951 and 2014 and with at least five consecutive years of data. Three final categories were drawn for each species depending on catch volume within each EEZ (e.g., present, maximum, and minimum EEZ's historical catch) and the year (e.g., year of maximum historical catch of the species within that EEZ) (Table \@ref(tab:CallTab2)) [@Kleisner:2011wn]. Accordingly, Category A represents fisheries that are registering increases in catch (“increasing”); Category B, species that have a constant catch rate (“constant”); and Category C, species that have registered declines in catch over the last 10 years (“decreasing”). Finally, we reported the predominant category over the time period considered.

@@

data.frame(Categories,
           Rules) %>% 
  kable("latex", 
        booktabs = T,
        caption = "Rules to determine the category of each transboundary species") %>%
  kable_styling(full_width = T) %>%
  column_spec(1,width = "3cm") %>% #wraps text in column2
  column_spec(2,width = "13cm") #%>% #wraps text in column2
  # write_csv("Table_2.csv")
```

@@

```{r Figure1, eval = T, fig.width=12, fig.height=9, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species and their contribution to global fisheries catch and revenue. a) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. b) Contribution of transboundary species to regional revenue (left) and catch (right). Regions classified according to the United Nations sub-regions. Points = mean +- sd. Revenue in 2010 real USD"}

@@

@@

Peru and Russia, having the two largest fisheries by total catch in the world [@FAO:2018KK], each responsible for over 5.8 million tonnes annually, on average, between 2005 and 2014  (Supplementary Fig. S1). Peru's catches consist mainly of Peruvian anchoveta (*Engraulis ringens*) and account for 79% of the national transboundary species production. Peru and Chile recently signed an agreement to work towards standardized stock assessments through coordinated management of the southern anchoveta stock [@UNDP:2016wp]. A similar management agreement was signed by Russia, Japan and the USA over shared Alaskan pollock (*Theragra chalcogramma*) in the Bering Sea in 1988 [@NOAA:gX6CtUTk]. Transboundary species also make large contributions to fisheries in Eastern Asia (one-way ANOVA; *Df*(16,158) = 2.265, *p* = 0.005, $\alpha$ = 0.05). China, the world's top fish producer [@FAO:2018KK], obtains one third (5.1 million tonnes) of its total catches from transboundary species, followed by Japan (1.8 million tonnes) and South Korea (1.06 million tonnes). Differences in the regional importance of transboundary fisheries can also be illustrated in terms of catch-revenue over the area ($km^2$) of the EEZ (Figure \@ref(fig:Figure2)). As an example, the aggregated EEZ area for all Northern European countries where transboundary species are present was 3.3 million $km^2$, the $6^{th}$ smallest of the 17 groups analyzed (Table \@ref(tab:TableS1)). However, it had the second highest average revenue (USD 26.1 thousand per $km^2$) and the highest average catch (19.9 tonnes per $km^2$) of transboundary species per EEZ area between 2005 and 2014. At the country level, countries from Western Europe accrued significantly more revenue from transboundary fisheries per $km^2$ than any other country (one-way ANOVA, *Df*(16,165) = 3.267, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.05; Figure \@ref(fig:Figure2)).

```{r Figure2, eval = T, fig.width=8, fig.height=6, echo=FALSE, fig.pos= "H", fig.cap=""}

@@

We determined the catch trend of each species within each EEZ, classifying them as increasing (Category A), constant (Category B) or decreasing (Category C) (Figure \@ref(fig:Figure3)). While previous studies have demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Our intention here is to examine where the catch trends of transboundary species differ from non-transboundary species between 2005 and 2010 relative to historic catch since 1951 (See methods). We found significant differences for all catch trend categories for transboundary species (one-way ANOVA, *Df*(2,459) = 47.94, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.001), and no significant differences in catch trend categories for discrete species (one-way ANOVA, *Df*(2,106) = 1.885, *p* = 0.157; $\alpha$ = 0.05). We also found significant differences in catch trends when directly comparing transboundary to discrete species categories (MANOVA, *Df*(2,459) = 19.001, *p* < 0.001). Overall, transboundary species only targeted by one country are generally less likely to have a decreasing catch trend compared to instances where the shared species is fished by neighboring countries (Table \@ref(tab:TableS2)).

```{r Figure3, eval = T, fig.width=12, fig.height=12, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species by catch category. Category A, increasing catch; Category B, constant catch; Category C, decreasing catch"}

img <- readPNG(here("./Figures/Fig3.png"))
grid.raster(img)
```

@@

```{r Figure4, eval = T, fig.width=12, fig.height=14, echo=FALSE, fig.pos= "H", fig.cap="Number of countries shared by transboundary species. Showing only species that share > 20 countries"}

@@

```{r supplements, child = 'supplements.Rmd', eval = T}
```

### 2020-10-05 Scripts/supplements.Rmd
Set manuscript for thesis

---
title: "Supplementary information for The transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes, Gabriel Reygondeau, Colette C.C. Wabnitz and William W.L. Cheung"
output: pdf_document
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
- \floatplacement{table}{H}
---

```{r setup, eval = T, echo=F, warning=F,message=F, results='hide'}

@@

# Supplementary Figures

```{r Call_Sup_1, eval = T, fig.width = 12, fig.height = 10, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species per EEZ and their contributions to countries’ EEZ catch. The number of species is displayed in the EEZ while catch (in thousand tonnes) is displayed in the land polygon. Figure created using R version 3.5.2. Land and EEZ shapefiles from Natural Earth (www.naturalearthdata.com) and Sea Around Us (www.seaaroundus.org), respectively."}

@@

The number of grid cells in which a species is present (*Area Index*) is expected to directly influence whether or not a species is considered transboundary. Therefore, we tested the sensitivity of our results to the *Area Index* by estimating the number of transboundary species along a gradient of values (Supplementary Fig. S2). As expected, using a more relaxed value (e.g. the species is present within 10% or more of the EEZ) will result in a greater number of transboundary species (median of 40 EEZs shared per species), while a more restrictive value (e.g., 50%) results in a median estimate of 4. The current analysis utilized a 25% *Area Index* threshold (Supplementary Fig. S2 - darker histogram).

```{r Call_Sup_4, eval = T, fig.width=6, fig.height=9, echo=FALSE, fig.pos= "H", fig.cap=" Histogram of number of transboundary species using different Area Index threshold values. The darker histogram represents the threshold (25%) used in the present analysis. Figure created using R version 3.5.2."}

@@

# Supplementary Tables

```{r Call_SupT_1, eval = T, echo = F, fig.pos= "H"}

read.csv("~/GitHub/FishForVisa/Tables/Table_S1.csv") %>% 
  clean_names() %>% 
  filter(sub_region != "Other") %>% 
  arrange(sub_region) %>% 
  rename("UN sub region" = sub_region,
         "Area of EEZs (Km^2)" = size_entity_sum) %>% 
  kable("latex",
        booktabs = T,
        caption= "Size of all the EEZs in each sub region determined by the United Nations") %>% 
  kable_styling(full_width = T)

@@

```{r Call_SupT_2, eval = T, echo = F, fig.pos= "H", message = F, warning = F}

@@

read.csv("~/GitHub/FishForVisa/Tables/Table_S2.csv") %>% 
  clean_names() %>% 
  mutate(
    "Mean shared" = paste(round(mean_value_x,2)," \u00B1 (",round(sd_value_x,2),")",sep=""),
    "Mean not-shared" = paste(round(mean_value_y,2)," \u00B1 (",round(sd_value_y,2),")",sep="")
  ) %>%
  select(1,8:9) %>%
  filter(observation != "No Category") %>%
  rename(Category = observation) %>%
  arrange(Category) %>% 
  kable("latex",
        booktabs = T,
        caption= "Average \u00B1 (standard deviation) number of shared and non-shared transboundary species per EEZ for each catch trend category")

@@

We report the average fishing revenue of 2005 to 2014 derived from the fishing activity within the global EEZs.

>>>>

---
output:
  bookdown::pdf_book:
    toc : no
    keep_tex: no
    number_sections: yes
---

```{r setup_sup, eval = T, echo=F, warning=F,message=F, results='hide'}

@@

```{r CallSup1, eval = T, fig.width = 12, fig.height = 10, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species per EEZ and their contributions to countries’ EEZ catch. The number of species is displayed in the EEZ while catch (in thousand tonnes) is displayed in the land polygon. Figure created using R version 3.5.2. Land and EEZ shapefiles from Natural Earth (www.naturalearthdata.com) and Sea Around Us (www.seaaroundus.org), respectively."}

@@

The number of grid cells in which a species is present (*Area Index*) is expected to directly influence whether or not a species is considered transboundary. Therefore, we tested the sensitivity of our results to the *Area Index* by estimating the number of transboundary species along a gradient of values (Figure \@(fig:CallSup2)). As expected, using a more relaxed value (e.g. the species is present within 10% or more of the EEZ) will result in a greater number of transboundary species (median of 40 EEZs shared per species), while a more restrictive value (e.g., 50%) results in a median estimate of 4. The current analysis utilized a 25% *Area Index* threshold (Figure \@(fig:CallSup2 - darker histogram).

```{r CallSup2, eval = T, fig.width=6, fig.height=9, echo=FALSE, fig.pos= "H", fig.cap=" Histogram of number of transboundary species using different Area Index threshold values. The darker histogram represents the threshold (25%) used in the present analysis. Figure created using R version 3.5.2."}

@@

```{r TableS1, eval = T, echo = F}

read.csv("../Tables/Table_S1.csv") %>% 
  clean_names() %>% 
  filter(sub_region != "Other") %>% 
  arrange(sub_region) %>% 
  rename("UN sub region" = sub_region,
         "Area of EEZs (Km^2)" = size_entity_sum) %>% 
  kable("latex",
        booktabs = T,
        caption= "Size of all the EEZs in each sub region determined by the United Nations") %>% 
  kable_styling(full_width = T)

@@

```{r TableS2, eval = T, echo = F, message = F, warning = F}

@@

read.csv("../Tables/Table_S2.csv") %>% 
  clean_names() %>% 
  mutate(
    "Mean shared" = paste(round(mean_value_x,2)," \u00B1 (",round(sd_value_x,2),")",sep=""),
    "Mean not-shared" = paste(round(mean_value_y,2)," \u00B1 (",round(sd_value_y,2),")",sep="")
  ) %>%
  select(1,8:9) %>%
  filter(observation != "No Category") %>%
  rename(Category = observation) %>%
  arrange(Category) %>% 
  kable("latex",
        booktabs = T,
        caption= "Average \u00B1 (standard deviation) number of shared and non-shared transboundary species per EEZ for each catch trend category")

@@
